COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2019-10-10 main.tex
Update figure 6 and explanation

>>>>

As is seen in figure \ref{fig:long_ph}, the relatively fast precession rates induced by the mean motion resonances overpower the secular forcing by Jupiter and randomize the orientation of orbits near the nominal resonance locations. Of course, this difference in longitude of perihelion of bodies inside and outside of resonance is only present in the EJ simulation because the Jupiter imposes zero forced eccentricity on the disk when it has a circular orbit. Figure \ref{fig:m_hist} provides another view of the difference in collision statistics inside and outside and of resonance. This shows the mean anomalies of bodies at the moment that they undergo a collision. The top row corresponds to the EJ simulation and the bottom rom corresponds to the CJ simulation. The left panels correspond to collisions that occur inside of the 2:1 MMR and the right panels are for collisions that occur with a semimajor axis between 2.7 and 2.8 AU, which is meant to well away from the influence of any resonances. The semimajor axis window size in the right hand panel is chosen so that both histograms contain the same total number of collisions.

The qualitative differences between these four panels can be explained in the following way: for $\approx$ 100 km bodies, gravitational focusing plays a significant role in setting the collision rate. In regions of the disk where the rms eccentricity is small (which corresponds to a small velocity dispersion), the collision rate should be enhanced (see equation \ref{eq:safronov}). Inside of resonances, planetesimals undergo large amplitude libration which pumps up the eccentricity dispersion. This effect is visible in figure \ref{fig:ae} for both the EJ and CJ simulation. This implies that the collision rate should be geometric, and not depend on the velocity dispersion near resonances. When this is true, collisions should preferentially occur near perihelion. The reason for this is that in the geometric limit, the collision probability scales with the ratio between the geometric cross section of the two planetesimals and the surface area of a sphere with radius $r$, where $r$ is the distance between the bodies and the central star. This ratio, and therefore the collision probability, is largest when $r$ is minimized (see \citet{2003AJ....125.2692L}). The fact both of the left-hand panels in figure \ref{fig:m_hist} show a distinct lack of collisions near aphelion is consistent with this explanation.

Note that this trend is not present in the right-hand panels of figure \ref{fig:m_hist}. In the case with Jupiter on a circular orbit (bottom right panel), there is a spike in the collision rate both at perihelion and aphelion. The perihelion spike is still due to the geometric argument presented in the previous paragraph. The aphelion spike is due to the effect of gravitational focusing. At aphelion, bodies achieve their minimum speed. This lowers the relative velocity between planetesimals, which according to equation \ref{eq:safronov} should enhance the collision rate. Note that the collision rate does not exhibit much dependence on mean anomaly outside of resonance in the case with Jupiter on an eccentric orbit (top right panel). This is due to the secular forcing by Jupiter, which aligns the longitude of perihelia of adjacent planetesimals. The result of this is that bodies maintain similar relative speeds over the course of an orbit, even though the orbit is eccentric. This lack of variation in the relative speeds makes collisions equally likely at any orbital phase.

\subsection{Comparison With Analytic Safronov Collision Rate}\label{sec:saf}
